---
title: "SimOps-Project"
output: html_document
date: "2024-04-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library need for Integer Programming
```{r}
# We will use the lpSolve package to solve our lp models.
# The `quietly = TRUE` option will not display the warning messages usually 
library(lpSolve, quietly = TRUE)
```


## Part I
Constructing a maximization integer program for quantity of bake goods for 
Ryan Hall only.
```{r}
# Specify the objective function for Ryan Hall model
ryan_obj_fun <- c(1, 1, 1, 1, 1, 1, 1)

# Set up constrains coefficients
ryan_lhs_constr <- matrix(c(1.1, .58, .58, 1.42, 2.46, 3.6, 1.42,
                     rep(0,4),1,0,1,
                     rep(1,4),0,1,0,
                     rep(1,7),
                     1,rep(0,4),-3,0,
                     rep(0,3), 1, rep(0,3),
                     0, 1, -1, rep(0,4),
                     1,rep(0,6),
                     0,1,rep(0,5),
                     rep(0,2),1, rep(0,4),
                     rep(0,3),1,rep(0,3),
                     rep(0,4),1,rep(0,2),
                     rep(0,5),1, 0,
                     rep(0,6),1), ncol = 7, byrow = TRUE)


# Set up constraint levels
ryan_const_RHS <- c(500, 30, 170, 200, 0, 60, 0, 80, rep(20,2), 40, rep(10,2), 20)

#Set up constrain direction
ryan_const_dir <- c("<=", 
               rep(">=",4),
               "<=",
               "=",
               rep(">=",7))

ryan_model <- lp(direction = "max",
                     objective.in = ryan_obj_fun,
                     const.mat = ryan_lhs_constr,
                     const.dir = ryan_const_dir,
                     const.rhs = ryan_const_RHS,
                     all.int = TRUE)
# Mix of bake goods product
ryan_model$solution
# Chocolate Chip cookies X1 = 80
# Sugar cookies X2 = 229
# Snicker doodle cookies X3 = 229
# Brownies X4 = 40
# Muffin Gluten Free (GF) X5 = 10
# Cinnamon rolls X6 = 10
# Chocolate Chip Cookies GF X7 = 20

# Total quantity made for Ryan Hall
ryan_model$objval
# Total = 618
```

## Part II
Improve the Model to reduce the overwhelming number of sugar and snicker diddle 
cookies. 
```{r}

# Specify the objective function for Ryan Hall improved product mix
improve_obj_fun <- c(1, 1, 1, 1, 1, 1, 1)

# Set up constrains coefficients
improve_lhs_constr <- matrix(c(1.1, .58, .58, 1.42, 2.46, 3.6, 1.42,
                     rep(0,4),1,0,1,
                     rep(1,4),0,1,0,
                     rep(1,7),
                     1,rep(0,4),-3,0,
                     rep(0,3), 1, rep(0,3),
                     0, 1, -1, rep(0,4),
                     1,rep(0,6),
                     0,1,rep(0,5),
                     rep(0,2),1, rep(0,4),
                     rep(0,3),1,rep(0,3),
                     rep(0,4),1,rep(0,2),
                     rep(0,5),1, 0,
                     rep(0,6),1,
                     -.25,.75,rep(-.25,5)), ncol = 7, byrow = TRUE)


# Set up constraint levels
improve_const_RHS <- c(500, 30, 170, 200, 0, 60, 0, 80, rep(20,2), 40, rep(10,2), 20, 0)

#Set up constrain direction
improve_const_dir <- c("<=", 
               rep(">=",4),
               "<=",
               "=",
               rep(">=",7),
               "<=")

improve_model <- lp(direction = "max",
                     objective.in = improve_obj_fun,
                     const.mat = improve_lhs_constr,
                     const.dir = improve_const_dir,
                     const.rhs = improve_const_RHS,
                     all.int = TRUE)

improve_model$solution
# Chocolate Chip cookies X1 = 181
# Sugar cookies X2 = 131
# Snicker doodle cookies X3 = 131
# Brownies X4 = 42
# Muffin Gluten Free (GF) X5 = 10
# Cinnamon rolls X6 = 10
# Chocolate Chip Cookies GF X7 = 20

# Total quantity made for Ryan Hall with a better mix
improve_model$objval
# Total = 525
```


## Part III
Constructing a maximization integer program for quantity of bake goods 
for Campus
```{r}
# Specify the objective function for all of campus
campus_obj_fun <- c(1, 1, 1, 1, 1, 1, 1)

# Set up constrains coefficients
campus_lhs_constr <- matrix(c(1.1, .58, .58, 1.42, 2.46, 3.6, 1.42,
                     rep(0,4),1,0,1,
                     rep(1,4),0,1,0,
                     rep(1,7),
                     1,rep(0,4),-3,0,
                     rep(0,3), 1, rep(0,3),
                     0, 1, -1, rep(0,4),
                     1,rep(0,6),
                     0,1,rep(0,5),
                     rep(0,2),1, rep(0,4),
                     rep(0,3),1,rep(0,3),
                     rep(0,4),1,rep(0,2),
                     rep(0,5),1, 0,
                     rep(0,6),1), ncol = 7, byrow = TRUE)


# Set up constraint levels
campus_const_RHS <- c(500, 45, 255, 300, 0, 72, 0, 120, rep(30,2), 60, rep(15,2), 30)

#Set up constrain direction
campus_const_dir <- c("<=", 
               rep(">=",4),
               "<=",
               "=",
               rep(">=",7))

campus_model <- lp(direction = "max",
                     objective.in = campus_obj_fun,
                     const.mat = campus_lhs_constr,
                     const.dir = campus_const_dir,
                     const.rhs = campus_const_RHS,
                     all.int = TRUE)

campus_model$solution
# Chocolate Chip cookies X1 = 120
# Sugar cookies X2 = 128
# Snicker doodle cookies X3 = 128
# Brownies X4 = 60
# Muffin Gluten Free (GF) X5 = 15
# Cinnamon rolls X6 = 15
# Chocolate Chip Cookies GF X7 = 30

# Total quantity made for Campus
campus_model$objval
# Total = 496

```

## Part IV
Constructing a maximization integer program for quantity of bake goods after 
inflation causes prices to rise by 18%
```{r}
# Specify the objective function for all of campus
inflation_obj_fun <- c(1, 1, 1, 1, 1, 1, 1)

# Set up constrains coefficients
inflation_lhs_constr <- matrix(c(1.1*1.18, .58*1.18, .58*1.18, 1.42*1.18, 2.46*1.18, 3.6*1.18, 1.42*1.18,
                     rep(0,4),1,0,1,
                     rep(1,4),0,1,0,
                     rep(1,7),
                     1,rep(0,4),-3,0,
                     rep(0,3), 1, rep(0,3),
                     0, 1, -1, rep(0,4),
                     1,rep(0,6),
                     0,1,rep(0,5),
                     rep(0,2),1, rep(0,4),
                     rep(0,3),1,rep(0,3),
                     rep(0,4),1,rep(0,2),
                     rep(0,5),1, 0,
                     rep(0,6),1), ncol = 7, byrow = TRUE)


# Set up constraint levels
inflation_const_RHS <- c(500, 45, 255, 300, 0, 72, 0, 120, rep(30,2), 60, rep(15,2), 30)

#Set up constrain direction
inflation_const_dir <- c("<=", 
               rep(">=",4),
               "<=",
               "=",
               rep(">=",7))

inflation_model <- lp(direction = "max",
                     objective.in = inflation_obj_fun,
                     const.mat = inflation_lhs_constr,
                     const.dir = inflation_const_dir,
                     const.rhs = inflation_const_RHS,
                     all.int = TRUE)

inflation_model$solution
# Chocolate Chip cookies X1 = 121
# Sugar cookies X2 = 62
# Snicker doodle cookies X3 = 62
# Brownies X4 = 60
# Muffin Gluten Free (GF) X5 = 15
# Cinnamon rolls X6 = 15
# Chocolate Chip Cookies GF X7 = 30

# Total quantity made for Campus after inflation
inflation_model$objval
# Total = 365
```

